\boldmath Primordial Black Hole Formation in a Dust Bouncing Model

### 2024-12-04 main.tex
Second refereee response.

\subsection{Wheeler-DeWitt Equation Solutions}

@@

Eliminating the time $t$ using \eqref{bohm_time} in \eqref{bohm_hubble}, we find
\begin{equation}
	\label{friedman_scale}
	\bar{H}^{2} = \frac{ 4 }{ 9t^{2}_{b}}\dpar
	{ \frac{ \bar{a}^{3}_{b} }{ \bar{a}^{3} } - \frac{\bar{a}_{b}^{6}}{ \bar{a}^{6} } }\, ,
\end{equation}
which is equivalent to the Friedmann equations
\begin{equation}\label{effective_friedmann}
	\bar{H}^{2} = \frac{\kappa}{ 3 }\bar\rho - \bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6},
\end{equation}
where $\bar\rho$ is the dust energy density. Also, from the derivative of
Eq.~\eqref{friedman_scale},
\begin{align}
	\dot{\bar{H}} = - \frac{\kappa}{2} \bar\rho + 3\bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6}.
\end{align}

In the above equations, the overdot represents the time derivative and there is an
additional term $-\bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6}$ when compared to the usual
Friedmann equation for a classical universe. In our model, the scale factor dynamics
resemble a typical Friedmann equation with a total energy density $\rho_{T} = \bar{\rho}
	+ \rho_{q},$ where $\bar{\rho}$ is the dust energy density, and $\rho_{q} = -\Omega
	\rho_{c} \bar{a}^{-6}$ denotes an effective negative energy density with an equation of
state $w_q = 1$ that accounts for quantum effects to avoid the primordial
singularity~\cite{vitenti2012large}. It is important to note that $\rho_{q}$ is not a
physical energy density but rather an effective one that emerges from the Bohmian
trajectory. In the perturbative analysis, the energy density will always be associated
only with $\bar{\rho}$. For large-scale factors $\bar{a} \gg \bar{a}_{B} $,
Eq.~\eqref{effective_friedmann} and Eq.~\eqref{bohm_scale_factor} reduce to the
classical dust dominated a flat FLRW universe. With this, we conclude the analysis of
our background quantum bouncing model and move to its associated perturbations. In the
next sections, we will consider the above quantum bouncing model with $w\approx
	10^{-10}$ and $\bar{a}(t_0)/a_B = 10^{35}$ where $t_0$ is the time during the
contracting phase where $H(t_0) = -H_0$ for $H_0 = 70 \, \text{km/s/Mpc}$. Moreover,
when solving the equations numerically, we use the exact expressions in terms of $w$
without approximation.

@@

We are mostly interested in the energy density perturbations that collapse to form PBHs.
These perturbations can be examined using the density contrast, defined as
\begin{align}
	\label{densitycon}
	\delta & \equiv \frac{\delta \rho}{\bar{\rho} + \bar{p}}
	,\end{align}
where $\delta\rho$ is the perturbation to the background energy density $\bar\rho$. The
density contrast provides a normalized measurement of the energy density perturbation
around the background matter density field. However, once again we are interested in its
gauge-invariant form
\begin{align}
	\label{deltarhoinvariant}
	\tilde{\delta\rho} & \equiv \delta\rho - \mathcal{V} \dot{\Bar{\rho}}
	,\end{align}
where $\mathcal{V}$ is the velocity perturbation of the fluid, which has its gauge
invariant form
\begin{align}
	\tilde{\mathcal{V}} & \equiv \mathcal{V} + \delta\sigma
	.\end{align}

@@

@@

The critical collapse model assumes the existence of a region of radius $r$ with an
overdensity $\delta$. To compute the probability of the existence of this overdense
region, we will need to compute the variance of our random variable $\delta$ and instead
of working with the original density contrast, we work with its filtered version
$\delta_r$, which is defined by
\begin{align}
	\delta_r(t) & =  \int d^{3}\textbf{x}^\prime \delta(t, \vec{x}^\prime) W_r\left(\vec{x} - \vec{x^\prime} \right)
	,\end{align}
where $r$ is a smoothing scale in comoving units related to the mean density by $M =
	\frac{4\pi r^3 \bar{\rho}}{3}$ and $W$ is the top-hat filter
\begin{align}
	\label{windowf}
	W_r(\vec{x}) & =W_r\left(|\vec{x}| = x\right) =
	\begin{cases}
		\frac{1}{4\pi r^3} & \text{if $x \leq r$} \\
		0                  & \text{otherwise}
	\end{cases}
	.\end{align}
The filter function is used to select a desired scale such that we would only analyze
collapsed objects that lie in this interval\footnote{Frequently we will usually refer to
	such scale in terms of the mass $M$ instead of the comoving radius $r$ that encloses
	this mass.}. In this work, we are using a simple top-hat windowed function for the
filter. We usually work on the Fourier space with comoving wave-number $k$, such that
the filter is given by
\begin{align}
	W_r(k) & = \frac{3}{(kr)^2}\left(\frac{\sin(kr)}{kr} - \cos(kr)\right) = \frac{3}{kr}j_1(kr)
	,\end{align}
where $j_1$ is the spherical Bessel function of the first kind. We will see that the
filter is necessary to analyze scales where the PBH formation is most likely to happen.

Coming back to $\delta_r$, we know from our quantum fields that the density contrast is
a random field following a Gaussian distribution with a zero mean. This is also
supported by significant research on the statistics of peaks in random Gaussian fields,
which can give rise to collapsed objects, as explored in
Ref.~\cite{Bardeen1986statistics}. Mathematically, the probability of a region with
radius $r$ having an overdensity $\delta_i$ is given by
\begin{align}
	\label{gaussdelta}
	P(\delta_i) & = \frac{1}{\sqrt{2\pi} \sigma_r} \exp(-\frac{\delta^2_i}{2 \sigma_r^2})
	.\end{align}
This Gaussian is completely defined by its variance
\begin{align}
	\label{eqsigmar}
	\sigma_r^2 & =\left< \delta_r^2(\vec{x}) \right> \equiv \sigma_r^2 = \frac{1}{2\pi^2} \int_0^\infty \frac{ \mathrm{~d}{k}}{k}\left[ P_{\tilde{\delta}}(k) \abs{W_r(k)}^2\right]
	,\end{align}
where $P_{\tilde{\delta}}(k)$ is given by Eq.~\eqref{eqspec} with the density contrast
modes. Unfortunately, for our bounce model, this integral does not converge. Before we
perform this calculation, we will analyze the critical threshold for PBH formation,
which will lead to scale constraints that will help us solve this problem.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Critical Threshold}

\label{critical_delta}
In inflationary models, the frozen super-Hubble density perturbations re-enter the
Hubble horizon at the end of the potential decay and collapse into a black hole if they
have values above the critical threshold~\cite{Martin2014}. Hence the Hubble horizon is
viewed as a characteristic scale for their formation and the threshold ($\delta_c$) is
obtained via the Minster-Sharp equations (see Ref.~\cite{Musco2019}). However, in the
bounce scenario, the perturbations constantly evolve in time and there is not only one
characteristic scale for the formation, since any perturbation above a critical
threshold may collapse as they are not frozen. Thus, we must carefully analyze how to
obtain this critical value starting with a local metric for the collapse.

In App.~\ref{appc}, we find the solutions for Einstein's equations of a critical
collapse represented with a local metric that depends on the local function $R(t,r)$,
which led to the LTB set of solutions
\begin{align}
	\label{ltb3a}
	R(\theta, r) & = \frac{r (1+\delta_{ini})}{\delta_{ini}}\sin^2\left(\frac{\theta}{2}\right),                                        \\
	\label{ltb4a}
	t(\theta,r ) & = t_1(r) + \frac{1 + \delta_{ini}}{2\bar{H}_{ini}\delta^{\frac{3}{2}}_{ini}}\left(\theta - \pi - \sin \theta \right)
	,\end{align}
such that we use the notation $\delta_{ini} = \delta(t_{ini})$ at the initial collapse
time. In this section, we wish to impose some constraints on those solutions to find the
exact point at which the perturbations will form a black hole. Consequently, this will
lead to a critical value for the density contrast. First of all, let us analyze the
parameter $\theta$. From Eq.~\eqref{sintheta},
\begin{align}
	\label{thetadef}
	\theta_{ini} & = \pm 2\arcsin(\sqrt{\frac{\delta_{ini}}{1+\delta_{ini}}})
	.\end{align}
To choose between the positive or negative value, we use our initial condition in
Eq.~\eqref{hinicond} and Eq.~\eqref{hdefltb} as a test, such that replacing
Eq.~\eqref{thetadef}
\begin{align}
	\left.\frac{\dot{R}}{R}\right|_{ini} & =  \left.\frac{\frac{\partial R}{\partial \theta}}{R \frac{\partial t}{\partial \theta}}\right|_{ini} = \pm \bar{H}_{ini}
	.\end{align}
From the above, we note that the positive sign in Eq.~\eqref{thetadef} leads to the
right definition of our initial conditions. If one chooses the negative sign, and as
$\bar{H}_{ini}$ is negative in the contracting phase, we would end up with an initially
expending patch. This feature can be seen in Fig.~\ref{theta_evol}, where
Eqs.~\eqref{ltb3a} and \eqref{ltb4a} were solved analytically with Wolfram
Mathematica~\cite{Mathematica}. The blue lines represent the positive choice that leads
to the LTB collapse model in a contracting universe, while the orange lines describe an
LTB collapse model with an initially expanding patch. On the left side of the figure,
for a larger value of $\delta_{ini}$, we see that both solutions are similar as the
orange plot has an insignificant initial expansion. For a smaller value of
$\delta_{ini}$ on the right, the model in orange first goes on an expansion phase
followed by the collapse, while the blue plot is already in a contracting phase and
collapses much earlier.

@@

From the right side of Eq.~\eqref{mprime}, we see that $m$ is related to the mass
density of a spherical volume. To analyze this quantity, we start by rewriting the
energy density from the spherical collapse metric as
\begin{align}
	\rho(t, r) & = \bar{\rho}(t)(1 + \delta(t,r))
	,\end{align}
such that $\delta$ is the density contrast $\delta = \frac{\rho(t,r) -
		\bar{\rho}(t)}{\bar{\rho}(t)}$ and $\bar{\rho}$ the background energy density.  We
consider that there is a small overdensity in our spherical system when compared to the
background density. We want to study a top-hat spherical collapse so that outside the
over-dense spherical shell characterized by a critical radius $r_c$, the energy density
is approximately the background density, i.e., ($r > r_c$)$\rightarrow$ $\rho \approx
	\bar{\rho}$. We can add this information to the density contrast by making the
redefinition
\begin{align}
	\delta(t, r) & \rightarrow \delta(t, r) \Theta(r-r_c)
\end{align}
where $\Theta$ is the Heaviside function. In this context, Eq.~\eqref{mprime} becomes
\begin{align}
	m(r) & = \int_0^r dr_1~ 4\pi GR^2R^\prime \left(\bar{\rho} + \delta \Theta(r-r_c)\right)
	.\end{align}

We still have freedom in our metric to define the local function $R$, which can be fixed
with initial conditions. The simplest choice would be that at initial collapse time
$t_{ini}$, $R(t_{ini},r) = r$, which implies that the 3-dimensional spherical shell is
initially at rest\footnote{Note that any other initial condition is valid and sufficient
	to completely define $m$.}. This leads to
\begin{align}
	\label{mr1}
	m(r) & = \frac{4\pi r^3G \bar{\rho}(t_{ini})}{3}\left(1 + \frac{3}{r^3}\int_0^r dr_1 (r_1)^2\delta(t_{ini}, r_1)\Theta(r-r_c)\right)
	.\end{align}
The integral represents the expected value of the density contrast inside a spherical
region with radius $r$ and $m$ gives the total mass inside this region. To evaluate it,
we need to assume that at the initial time, our spherical metric possesses the same
symmetries as the background and thus it is homogeneous. This means that the density
contrast has a uniform distribution and does not depend on position at
$t_{ini}$\footnote{This is only true at $t_{ini}$. At later times, the perturbation
	evolves and depends on the radial position.}. Consequently,
\begin{align}
	\label{deltaini}
	\delta(t_{ini}, r) & = \delta(t_{ini}) \equiv  \delta_{ini}.
\end{align}
We are representing any variable $V$ at initial time $t_{ini}$ as $V(t_{ini})\equiv
	V_{ini}$ to simplify the notation. Plugging Eq.~\eqref{deltaini} in \eqref{mr1} leads to
\begin{align}
	\label{pbhmassmetric}
	m (r) & =  \left\{\begin{array}{ll}
		                  M\left(\frac{r^3}{r_c^3}\right)                                       & r \leq r_c \\
		                  M+\frac{M}{1+\delta_{ini}}\left(\frac{r^3}{r_{\mathrm{c}}^3}-1\right) & r > r_c
	                  \end{array}\right.,
\end{align}
such that we rewrote the terms as a function of
\begin{align}
	M & =m(r_c)=\frac{4\pi G \bar{\rho}_{ini}r^3_c}{3}\left(1 + \delta_{ini}\right)
	.\end{align}

>>>>

\subsection{Wheeler-DeWitt Equation Solutions}
\label{sec:wdw}

@@

Eliminating the time $t$ using \eqref{bohm_time} in \eqref{bohm_hubble}, we find
\begin{equation}
	\label{friedman_scale}
	\bar{H}^{2} = \frac{ 4 }{ 9t^{2}_{b}}\dpar
	{ \frac{ \bar{a}^{3}_{b} }{ \bar{a}^{3} } - \frac{\bar{a}_{b}^{6}}{ \bar{a}^{6} } }\, ,
\end{equation}
which is equivalent to the Friedmann equations
\begin{equation}\label{effective_friedmann}
	\bar{H}^{2} = \frac{\kappa}{ 3 }\bar\rho - \bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6},
\end{equation}
where $\bar\rho$ is the dust energy density. Also, from the derivative of
Eq.~\eqref{friedman_scale},
\begin{align}\label{effective_friedmann2}
	\dot{\bar{H}} = - \frac{\kappa}{2} \bar\rho + 3\bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6}.
\end{align}

In the above equations, the overdot represents the time derivative and there is an
additional term $-\bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6}$ when compared to the usual
Friedmann equation for a classical universe. In our model, the scale factor dynamics
resemble a typical Friedmann equation with a total energy density $\rho_{T} = \bar{\rho}
	+ \rho_{q},$ where $\bar{\rho}$ is the dust energy density, and $\rho_{q} = -\Omega
	\rho_{c} \bar{a}^{-6}$ denotes an effective negative energy density with an equation of
state $w_q = 1$ that accounts for quantum effects to avoid the primordial
singularity~\cite{vitenti2012large}. It is important to note that $\rho_{q}$ is not a
physical energy density but rather an effective one that emerges from the Bohmian
trajectory.

	{
		\color{red} As a final remark, it is evident that, in the classical theory, one
		could interchangeably use either $\bar{H}^2$ or $\kappa \bar{\rho}/3$, as they are
		related by the classical Friedmann equation. However, in the quantum framework,
		$\bar{\rho}$ retains the same functional dependence on the scale factor, whereas
		$\bar{H}^2$ acquires an additional term due to quantum effects absent in the
		classical theory. Consequently, when deriving the Lagrangian for perturbations, it
		is crucial to avoid relying on the classical Friedmann equations. This approach was
		explicitly followed in Ref.~\cite{fluidgeral}, where the authors computed
		perturbations up to second order without invoking any classical background
		equations.

		To compute perturbations in a Bohmian trajectory, one employs
		Eqs.~\eqref{effective_friedmann} and \eqref{effective_friedmann2} to determine the
		Hubble function and its time derivative, while the energy density $\bar{\rho}
			\propto \bar{a}^{-3}$ (with $\bar{a}$ given by Eq.~\eqref{bohm_scale_factor}) is
		used for calculating energy density perturbations. This ensures, for instance, that
		terms such as $1/(\bar{\rho} + \bar{p})$ remain well-defined even when $\bar{H}^2$
		or $\dot{\bar{H}}$ pass through zero.
	}

With this, we conclude the analysis of our background quantum bouncing model and move to
its associated perturbations. In the next sections, we will consider the above quantum
bouncing model with $w\approx 10^{-10}$ and $\bar{a}(t_0)/a_B = 10^{35}$ where $t_0$ is
the time during the contracting phase where $H(t_0) = -H_0$ for $H_0 = 70 \,
	\text{km/s/Mpc}$. Moreover, when solving the equations numerically, we use the exact
expressions in terms of $w$ without approximation.

@@

We are mostly interested in the energy density perturbations that collapse to form PBHs.
These perturbations can be examined using the density contrast, defined as
\begin{align}
	\label{densitycon}
	\delta & \equiv \frac{\delta \rho}{\bar{\rho} + \bar{p}},
\end{align}
where $\delta\rho$ is the perturbation to the background energy density $\bar\rho$.
	{\color{red} Note that the density contrast defined above appears in this form
		when the second-order perturbations Lagrangian is derived. This means
		that as long as $\bar{\rho} + \bar{p} > 0$, the density contrast is well-defined.
		As discussed in Sec.~\ref{sec:wdw}, when computing the perturbations,
		we use the Bohmian trajectory to determine the background quantities,
		which ensures that $\bar{\rho} + \bar{p} > 0$ for all times. }

The density contrast provides a normalized measurement of the energy density
perturbation around the background matter density field. However, once again we are
interested in its gauge-invariant form
\begin{align}
	\label{deltarhoinvariant}
	\tilde{\delta\rho} & \equiv \delta\rho - \mathcal{V} \dot{\Bar{\rho}}
	,\end{align}
where $\mathcal{V}$ is the velocity perturbation of the fluid, which has its gauge
invariant form
\begin{align}
	\tilde{\mathcal{V}} & \equiv \mathcal{V} + \delta\sigma
	.\end{align}

@@

{\color{red} Before we proceed, we must emphasize that we are considering the formation
of PBHs in a classical context. We are using the quantum fields to describe the initial
conditions of the perturbations, but the contracting phase is amplifying their
amplitudes in a similar fashion to an inflationary phase. Such assumptions, particularly
regarding the quantum-to-classical transition, are commonplace in the literature,
especially in inflationary models. For example, in Ref.~\cite{Q2C}, the authors discuss
the quantum-to-classical transition in the context of the Bohmian-De Broglie
interpretation, showing that the quantum potential is negligible when the perturbations
are amplified by the inflationary phase.

In this work, we are not discussing the quantum-to-classical transition but are assuming
that the perturbations are classical when amplified by the contracting phase. Moreover,
when approaching the bounce time, the background itself is not a classical solution but
a quantum one. Thus, in order to study the collapse of the perturbations at this phase,
one must consider the quantum effects of the background. However, it is important to
note that PBHs with masses within the observable ranges are formed during the
contracting phase, where it is safe to assume that the initial quantum fields have
already transitioned to classical behavior and are far removed from the bounce phase.}

@@

The critical collapse model assumes the existence of a region of radius $r$ with an
overdensity $\delta$. To compute the probability of the existence of this overdense
region, we will need to compute the variance of our random variable $\delta$ and instead
of working with the original density contrast, we work with its filtered version
$\delta_r$, which is defined by
\begin{align}
	\delta_r(t) & =  \int d^{3}\textbf{x}^\prime \delta(t, \vec{x}^\prime) W_r\left(\vec{x} - \vec{x^\prime} \right)
	,\end{align}
where $r$ is a smoothing scale in comoving units related to the mean density by $M =
	\frac{4\pi r^3 \bar{\rho}}{3}$ and $W$ is the top-hat filter
\begin{align}
	\label{windowf}
	W_r(\vec{x}) & =W_r\left(|\vec{x}| = x\right) =
	\begin{cases}
		\frac{1}{4\pi r^3} & \text{if $x \leq r$} \\
		0                  & \text{otherwise}
	\end{cases}
	.\end{align}
The filter function is used to select a desired scale such that we would only analyze
collapsed objects that lie in this interval\footnote{Frequently we will usually refer to
	such scale in terms of the mass $M$ instead of the comoving radius $r$ that encloses
	this mass.}. In this work, we are using a simple top-hat windowed function for the
filter. We usually work on the Fourier space with comoving wave-number $k$, such that
the filter is given by
\begin{align}
	W_r(k) & = \frac{3}{(kr)^2}\left(\frac{\sin(kr)}{kr} - \cos(kr)\right) = \frac{3}{kr}j_1(kr),
\end{align}
where $j_1$ is the spherical Bessel function of the first kind. We will see that the
filter is necessary to analyze scales where the PBH formation is most likely to happen.

Coming back to $\delta_r$, we know from our quantum fields that the density contrast is
a random field following a Gaussian distribution with a zero mean. This is also
supported by significant research on the statistics of peaks in random Gaussian fields,
which can give rise to collapsed objects, as explored in
Ref.~\cite{Bardeen1986statistics}. Mathematically, the probability of a region with
radius $r$ having an overdensity $\delta_i$ is given by
\begin{align}
	\label{gaussdelta}
	P(\delta_i) & = \frac{1}{\sqrt{2\pi} \sigma_r} \exp(-\frac{\delta^2_i}{2 \sigma_r^2}).
\end{align}
This Gaussian is completely defined by its variance
\begin{align}
	\label{eqsigmar}
	\sigma_r^2 & =\left< \delta_r^2(\vec{x}) \right> \equiv \sigma_r^2 = \frac{1}{2\pi^2} \int_0^\infty \frac{ \mathrm{~d}{k}}{k}\left[ P_{\tilde{\delta}}(k) \abs{W_r(k)}^2\right],
\end{align}
where $P_{\tilde{\delta}}(k)$ is given by Eq.~\eqref{eqspec} with the density contrast
modes. Unfortunately, for our bounce model, this integral does not converge. Before we
perform this calculation, we will analyze the critical threshold for PBH formation,
which will lead to scale constraints that will help us solve this problem.

\subsection{Critical Threshold}

\label{critical_delta}
In inflationary models, the frozen super-Hubble density perturbations re-enter the
Hubble horizon at the end of the potential decay and collapse into a black hole if they
have values above the critical threshold~\cite{Martin2014}. Hence the Hubble horizon is
viewed as a characteristic scale for their formation and the threshold ($\delta_c$) is
obtained via the Misner-Sharp equations (see Ref.~\cite{Musco2019}). However, in the
bounce scenario, the perturbations constantly evolve in time and there is not only one
characteristic scale for the formation, since any perturbation above a critical
threshold may collapse as they are not frozen. Thus, we must carefully analyze how to
obtain this critical value starting with a local metric for the collapse.

	{\color{red} It is important to note that we are assuming the perturbations are
		classical at this stage. Otherwise, it would be necessary to consider a quantum
		model for the collapse, which would require developing a quantum Misner-Sharp
		framework with initial conditions derived from the quantum fields. This is a highly
		complex problem that remains largely unexplored in the literature.}

In App.~\ref{appc}, we find the solutions for Einstein's equations of a critical
collapse represented with a local metric that depends on the local function $R(t,r)$,
which led to the Lemaître-Tolman-Bondi (LTB) set of solutions
\begin{align}
	\label{ltb3a}
	R(\theta, r) & = \frac{r (1+\delta_{ini})}{\delta_{ini}}\sin^2\left(\frac{\theta}{2}\right),                                        \\
	\label{ltb4a}
	t(\theta,r ) & = t_1(r) + \frac{1 + \delta_{ini}}{2\bar{H}_{ini}\delta^{\frac{3}{2}}_{ini}}\left(\theta - \pi - \sin \theta \right)
	,\end{align}
such that we use the notation $\delta_{ini} = \delta(t_{ini})$ at the initial collapse
time. In this section, we wish to impose some constraints on those solutions to find the
exact point at which the perturbations will form a black hole. Consequently, this will
lead to a critical value for the density contrast. First of all, let us analyze the
parameter $\theta$. From Eq.~\eqref{sintheta},
\begin{align}
	\label{thetadef}
	\theta_{ini} & = \pm 2\arcsin(\sqrt{\frac{\delta_{ini}}{1+\delta_{ini}}})
	.\end{align}
To choose between the positive or negative value, we use our initial condition in
Eq.~\eqref{hinicond} and Eq.~\eqref{hdefltb} as a test, such that replacing
Eq.~\eqref{thetadef}
\begin{align}
	\left.\frac{\dot{R}}{R}\right|_{ini} & =  \left.\frac{\frac{\partial R}{\partial \theta}}{R \frac{\partial t}{\partial \theta}}\right|_{ini} = \pm \bar{H}_{ini}.
\end{align}
From the above, we note that the positive sign in Eq.~\eqref{thetadef} leads to the
right definition of our initial conditions. If one chooses the negative sign, and as
$\bar{H}_{ini}$ is negative in the contracting phase, we would end up with an initially
expending patch. This feature can be seen in Fig.~\ref{theta_evol}, where
Eqs.~\eqref{ltb3a} and \eqref{ltb4a} were solved analytically with Wolfram
Mathematica~\cite{Mathematica}. The blue lines represent the positive choice that leads
to the LTB collapse model in a contracting universe, while the orange lines describe an
LTB collapse model with an initially expanding patch. On the left side of the figure,
for a larger value of $\delta_{ini}$, we see that both solutions are similar as the
orange plot has an insignificant initial expansion. For a smaller value of
$\delta_{ini}$ on the right, the model in orange first goes on an expansion phase
followed by the collapse, while the blue plot is already in a contracting phase and
collapses much earlier.

@@

From the right side of Eq.~\eqref{mprime}, we see that $m$ is related to the mass
density of a spherical volume. To analyze this quantity, we start by rewriting the
energy density from the spherical collapse metric as
\begin{align}
	\rho(t, r) & = \bar{\rho}(t)(1 + \delta(t,r))
	,\end{align}
such that $\delta$ is the density contrast $\delta = \frac{\rho(t,r) -
		\bar{\rho}(t)}{\bar{\rho}(t)}$ and $\bar{\rho}$ the background energy density,
{\color{red} note that this is the same density contrast as in
		Eq.~\eqref{densitycon} but using the approximation $w \ll 1$ for cold dark
		matter.} We consider that there is a small overdensity in our spherical system
when compared to the background density. We want to study a top-hat spherical
collapse so that outside the over-dense spherical shell characterized by a
critical radius $r_c$, the energy density is approximately the background
density, i.e., ($r > r_c$)$\rightarrow$ $\rho \approx \bar{\rho}$. We can add
this information to the density contrast by making the redefinition
\begin{align}
	\delta(t, r) & \rightarrow \delta(t, r) \Theta(r-r_c)
\end{align}
where $\Theta$ is the Heaviside function. In this context, Eq.~\eqref{mprime} becomes
\begin{align}
	m(r) & = \int_0^r dr_1~ 4\pi GR^2R^\prime \left(\bar{\rho} + \delta \Theta(r-r_c)\right).
\end{align}

We still have freedom in our metric to define the local function $R$, which can be fixed
with initial conditions. The simplest choice would be that at initial collapse time
$t_{ini}$, $R(t_{ini},r) = r$, which implies that the 3-dimensional spherical shell is
initially at rest\footnote{Note that any other initial condition is valid and sufficient
	to completely define $m$.}. This leads to
\begin{align}
	\label{mr1}
	m(r) & = \frac{4\pi r^3G \bar{\rho}(t_{ini})}{3}\left(1 + \frac{3}{r^3}\int_0^r dr_1 (r_1)^2\delta(t_{ini}, r_1)\Theta(r-r_c)\right).
\end{align}
The integral represents the expected value of the density contrast inside a spherical
region with radius $r$ and $m$ gives the total mass inside this region. To evaluate it,
we need to assume that at the initial time, our spherical metric possesses the same
symmetries as the background and thus it is homogeneous. This means that the density
contrast has a uniform distribution and does not depend on position at
$t_{ini}$\footnote{This is only true at $t_{ini}$. At later times, the perturbation
	evolves and depends on the radial position.}. Consequently,
\begin{align}
	\label{deltaini}
	\delta(t_{ini}, r) & = \delta(t_{ini}) \equiv  \delta_{ini}.
\end{align}
We are representing any variable $V$ at initial time $t_{ini}$ as $V(t_{ini})\equiv
	V_{ini}$ to simplify the notation. Plugging Eq.~\eqref{deltaini} in \eqref{mr1} leads to
\begin{align}
	\label{pbhmassmetric}
	m (r) & =  \left\{\begin{array}{ll}
		                  M\left(\frac{r^3}{r_c^3}\right)                                       & r \leq r_c \\
		                  M+\frac{M}{1+\delta_{ini}}\left(\frac{r^3}{r_{\mathrm{c}}^3}-1\right) & r > r_c
	                  \end{array}\right.,
\end{align}
such that we rewrote the terms as a function of
\begin{align}
	M & =m(r_c)=\frac{4\pi G \bar{\rho}_{ini}r^3_c}{3}\left(1 + \delta_{ini}\right)
	.\end{align}
